---
title: "lab6"
author: "Kyle Reardon"
date: "10/27/2019"
output: 
  html_document:
    highlight: pygments
  pdf_document:
    highlight: kate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 6.5,
                      fig.height = 8)
```

## R Markdown

The dataset appears to have 8 different selections for race/ethnicity, so I am not sure how to limit to just the four options in the PDF plot without knowing how each option is coded.

```{r load-data, echo = FALSE}

library(tidyverse)
library(here)
library(rio)
library(janitor)
library(tinytex)

eclsk <- import(here("data", "ecls-k_samp.sav")) %>% 
  characterize() %>% 
  clean_names()

unique(eclsk$ethnic)

eclsk %>% 
  filter(ethnic == "WHITE, NON-HISPANIC" | ethnic == "BLACK OR AFRICAN AMERICAN, NON-HISPANIC" | ethnic == "HISPANIC, RACE SPECIFIED" | ethnic == "HISPANIC, RACE NOT SPECIFIED") %>% 
  ggplot(aes(t1mscale, t2mscale)) +
  geom_point() +
  geom_smooth(aes(color = school_type), method = lm) +
  facet_wrap(~ethnic)

```

## Questions 4 and 5

```{r}
t1mean <- mean(eclsk$T1MSCALE, na.rm = TRUE)
t1sd <- sd(eclsk$T1MSCALE, na.rm = TRUE)

t2mean <- mean(eclsk$T2MSCALE, na.rm = TRUE)
t2sd <- sd(eclsk$T2MSCALE, na.rm = TRUE)


```

For time one, the mean is `r t1mean` and the standard deviation is `r t1sd`. For time two, the mean is `r t2mean` and the standard devisation is `r t2sd`. The average gain is `r t2mean - t1mean`.

## Reading in data

The most efficient way to load data is using the rio and here packages. Using the here package loads data from the working directory. This ensures that R is pulling from the right source throughout your code as opposed to having to specify each time.